Paper on the carbonate chemistry at Ny-Ålesund

### 2020-03-19 paper.Rmd
temp sal analyses

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
#library(robfilter)
library(seacarb)
library(gridExtra)
library(lubridate)
#library(lmtest)
#library(grid)
library(viridis)
#library(dygraphs)
require("knitr")
library("lmodel2")
#library(captioner)
#library(xts)
#library(seismicRoll)
library(scales)
library(plotly)
library(htmlwidgets)
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = normalizePath("/Users/gattuso/not_synced_git/AWIPEV-CO2-git"))

@@

```{r read and clean data, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- as_tibble(
  #readRDS(file = paste0(path_data, "fb_data/d_all.rds"))
  readRDS(file = paste0(path_data, "fb_data/d_all_copy.rds"))
) %>%
  filter(sal_fb > 28) # remove salinity data lower than 28
```

@@

```{r water masses, fig.width=15, fig.height=15, echo=FALSE, message= FALSE}
# One tries to identify water masses coming from the Atlantic or freshwater
temp_sal <- d_all %>%
  ggplot(aes(x = sal_fb, y = temp_insitu, col = month), na.rm=TRUE) +
  geom_point(colour="black", na.rm=TRUE, size=size_point) + 
  labs(title="Temperature vs salinity SBE45",x="Salinity", y="Temp. (°C)") +
  coord_fixed() +
  Mytheme(size_labs = 8) #+
  #theme(axis.text.x=element_blank())
#  +annotation_custom(grob_A)
print(temp_sal)

den <- seacarb::rho(S = d_all$sal_fb , T = d_all$temp_insitu, P = 1)
d_all <- d_all %>%
  dplyr::mutate(den = den)
ts_den <- d_all %>%
  ggplot(aes(x = datetime, y = den), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="black", na.rm=TRUE, size=size_point) + 
  labs(title="Density",x="Time", y="Density") +
  #coord_fixed() +
  Mytheme(size_labs = 8)
print(ts_den)
```

>>>>

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
#library(robfilter)
library(seacarb)
library(gridExtra)
library(lubridate)
#library(lmtest)
#library(grid)
library(viridis)
#library(dygraphs)
require("knitr")
library("lmodel2")
#library(captioner)
#library(xts)
#library(seismicRoll)
library(scales)
library(plotly)
library(htmlwidgets)
library(tsibble)
if (!require("directlabels")) install.packages("directlabels")
library("directlabels")
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = normalizePath("/Users/gattuso/not_synced_git/AWIPEV-CO2-git"))

@@

```{r read and clean data, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- as_tibble(readRDS(file = paste0(path_data, "fb_data/d_all_copy.rds"))) %>% 
  dplyr::filter(sal_fb > 28) %>% # remove salinity data lower than 28
  dplyr::mutate(year_month = format(datetime, '%Y-%m'))
```

@@

#Temperature vs salinity
We explore the use of a TS diagram to investigate changes in water sources. Now need to calculate slopes and r2, and see whether some months have a consistently different behaviour.

```{r temperature vs salinity plots, fig.width=8, fig.height=8, echo=FALSE, message= FALSE}
# One tries to identify water masses coming from the Atlantic or freshwater
s <- c(seq(30.4, 35.3, length.out = 100)) # pCO2
t <- seq(-2, 8, length.out = 100) #temperature throughout the year
grid <- as_tibble(expand.grid(s=s, t=t))
grid$z <- rho(S = grid$s, T = grid$t, P = 1)

d <- d_all %>%
  dplyr::filter(year < 2020) %>%
  dplyr::mutate(month=month(datetime, label = TRUE, abbr = TRUE)) # abbreviated months rather than digital months
  #ggplot(aes(x = sal_fb, y = temp_insitu), na.rm=TRUE) +
temp_sal <-  ggplot() +
  geom_point(data = d, aes(x = sal_fb, y = temp_insitu, fill=month), na.rm=TRUE,
             size=size_point, show.legend=FALSE) + 
  labs(title="Temperature vs salinity, blue lines show density", x="Salinity", y="Temp. (°C)") +
  facet_wrap(~ year, ncol = 3) +
  stat_smooth(data = d, aes(x = sal_fb, y = temp_insitu, color=month),
              method=lm, na.rm=TRUE, se=FALSE) +
  scale_color_viridis(name="Month", discrete=TRUE, option="plasma") +
  geom_contour(data = grid, aes(x = s, y = t, z = z), 
               stat = "contour", binwidth=1, show.legend=FALSE) +
  #stat_contour(aes(colour=..level..), breaks=seq(1025, 1028, by=1)) +
  coord_fixed() +
  Mytheme(size_labs = 10)
ggsave(temp_sal,filename = paste0(path_fig, "temp_sal.png"), width = 20, height = 20 ,units="cm")
temp_sal_py <- ggplotly(temp_sal,dynamicTicks = TRUE)
filename.html <- paste0(path_fig, "temp_sal_py.html")
saveWidget(temp_sal_py, filename.html, selfcontained = T, libdir = "lib")
```

```{r temperature vs salinity regressions, fig.width=8, fig.height=8, echo=FALSE, message= FALSE}
# AT as a function of salinity for every year
j <- 0
reg <- data.frame()
for (i in 1:length(unique(d$year_month))) { #one loops through all years
  j <- j + 1
  ym <- unique(d$year_month)[i]
  d_reg <- filter(d, year_month==ym)
  if (length(!is.na(d_reg$temp_fb)) > 2) { #there is no temp_insitu in 2016-08, hence skip. Not needed if one uses temp_fb
  lms <- summary(lm(data=d_reg, sal_fb ~ temp_fb))
  # calculate probability of fstatistic because it cannot be extracted from lms above
  # see http://stats.stackexchange.com/questions/92824/extracting-the-model-p-value-for-a-multiple-regression-in-r
  # slope returned by lm is per second
  prob <- pf(lms$fstatistic[1],lms$fstatistic[2],lms$fstatistic[3],lower.tail=FALSE)
  reg <- rbind(reg, as.numeric(c(lms$coefficients[2,1], lms$coefficients[2,2], 
                                 lms$coefficients[2,4], lms$coefficients[1,1], lms$coefficients[1,2],
                                 lms$coefficients[1,4], lms$fstatistic[1], lms$fstatistic[3], lms$r.squared, prob)))
  }
}
names(reg) <- c("Slope", "SE Slope", "P Slope","Intercept","SE int.","P int.", "F", "df",
                "R2","P value")
row.names(reg) <- as.character(unique(d$year_month))
#transform in latex table and save
tab_reg_temp_sal <- reg
tmp_reg_temp_sal <- reg %>% # to enable plotting below
  dplyr::mutate(date = as.Date(paste0(row.names(reg), "-01")),
                month = month(date, label=TRUE, abbr = TRUE))
#plot
fig_reg_temp_sal <- tmp_reg_temp_sal %>%
  ggplot() +
  geom_point(aes(x=date, y=Slope, color=month), size=5) +
  scale_color_viridis(name="Month", discrete = TRUE) +
  Mytheme(size_labs = 14)
fig_reg_temp_sal
```
kable(tab_fig_dyf_surf_date_AT)
print(dyf_surf_salinity_AT)

```{r density, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
# d_all <- d_all %>%
#   dplyr::mutate(den = den)
# ts_den <- d_all %>%
#   ggplot(aes(x = datetime, y = den), na.rm=TRUE) +
#   scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
#   geom_point(colour="black", na.rm=TRUE, size=size_point) + 
#   labs(title="Density",x="Time", y="Density") +
#   #coord_fixed() +
#   Mytheme(size_labs = 14)
# print(ts_den)
```

```{r pangaea data, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
data_2015 <- pg_data(doi="10.1594/PANGAEA.896771", overwrite = TRUE)
data_2016 <- pg_data(doi="10.1594/PANGAEA.896770", overwrite = TRUE)
data_2017 <- pg_data(doi="10.1594/PANGAEA.896170", overwrite = TRUE)
data_2018 <- pg_data(doi="10.1594/PANGAEA.897349", overwrite = TRUE)
tmp <- rbind(data_2015[[1]]$data, data_2016[[1]]$data, data_2017[[1]]$data, data_2018[[1]]$data)
tmp <- tmp[, c(1,8)] # select date time and salinity
tmp$`Date/Time` <- as.POSIXct(strptime(tmp$`Date/Time`, "%Y-%m-%dT%H:%M", tz = "UTC"))

all_data <- tmp %>%
  dplyr::filter(!are_duplicated(tmp)) %>%
  dplyr::rename(datetime = `Date/Time`,
                sal = `Sal ([PSU])`) %>%
  as_tsibble(key = NULL, index = datetime) %>%
  fill_gaps()

all_data %>%
  ggplot() +
  geom_point(aes(x=datetime, y=sal), na.rm = TRUE)
```
